Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-06-13 thesis/abstract.tex
abstract edits

\vskip.35in
{\large Andrew J. Emerick} \\
\vskip.35in
\end{center}
Motivated by the desire to investigate two of the largest outstanding problems in galactic evolution -- stellar feedback and galactic chemical evolution --
%In this \thesis
we develop the first set of galaxy-scale simulations that simultaneously follow star formation with individual stars and the multi-channel stellar feedback and multi-element metal yields associated with each star. We developed these simulations to probe the way in which stellar feedback, including stellar winds, stellar radiation, and supernovae, couples to the interstellar medium (ISM), regulates star formation, and drives outflows in dwarf galaxies. We follow the evolution of the individual metal yields associated with these stars in order to trace how metals mix within the ISM and are ejected into the circumgalactic and intergalactic media (CGM, IGM) through outflows. This study is directed with the ultimate goal of leveraging the ever increasing quality of stellar abundance measurements within our own Milky Way galaxy and in nearby dwarf galaxies to understand galactic evolution.

Our simulations follow the evolution of an idealized, isolated low mass dwarf galaxy ($M_{\rm vir} \sim 10^{9}$~M$_{\odot}$) over $\sim$ 500 Myr timescales using the adaptive mesh refinement hydrodynamics code \textsc{Enzo}. We implemented a new star formation routine which deposits stars individually from 1~M$_{\odot}$ to 100~M$_{\odot}$. Using tabulated stellar properties, we follow the stellar feedback from each star. For massive stars ($M_* > 8$~M$_{\odot}$) we follow their stellar winds, ionizing radiation (using an adaptive ray-tracing radiative transfer method), the FUV radiation which leads to photoelectric heating of dust grains, Lyman-Werner radiation, which leads to H$_2$ dissociation, and core collapse supernovae. In addition, we follow the asymptotic giant branch (AGB) winds of low-mass stars ($M_* < 8$~M$_{\odot}$) and Type Ia supernovae. We investigate how this detailed model for stellar feedback drives the evolution of low mass galaxies. We find agreement with previous studies that these low mass dwarf galaxies exhibit bursty, irregular star formation histories with significant feedback-driven winds.

Using these simulations, we investigate the role that stellar radiation feedback plays in the evolution of low mass galaxies.
% a fiducial, full-physics simulation, second, a simulation with ionizing radiation restricted to the region immediately around massive star particles, and third, a simulation with no ionizing radiation feedback. I
In the regime of low mass dwarf galaxies, we find that the local effects of stellar radiation (within $\sim$ 10~pc of the massive, ionizing source star) act to regulate star formation by rapidly destroying cold, dense gas around newly formed stars. For the first time, we find that the long-range radiation effects far from the birth sites are vital for carving channels of diffuse gas in the ISM which dramatically increase the effect of supernovae. We find this effect is necessary to drive strong winds with significant mass loading factors and has a significant impact on the metal content of the ISM.

>>>>

\vskip.35in
{\large Andrew J. Emerick} \\
\vskip.35in
\end{center}
Motivated by the desire to investigate two of the largest outstanding problems in galactic evolution -- stellar feedback and galactic chemical evolution --
%In this \thesis
we develop the first set of galaxy-scale simulations that simultaneously follow star formation with individual stars and their associated multi-channel stellar feedback and multi-element metal yields. We developed these simulations to probe the way in which stellar feedback, including stellar winds, stellar radiation, and supernovae, couples to the interstellar medium (ISM), regulates star formation, and drives outflows in dwarf galaxies. We follow the evolution of the individual metal yields associated with these stars in order to trace how metals mix within the ISM and are ejected into the circumgalactic and intergalactic media (CGM, IGM) through outflows. This study is directed with the ultimate goal of leveraging the ever increasing quality of stellar abundance measurements within our own Milky Way galaxy and in nearby dwarf galaxies to understand galactic evolution.

Our simulations follow the evolution of an idealized, isolated, low mass dwarf galaxy ($M_{\rm vir} \sim 10^{9}$~M$_{\odot}$) for $\sim$ 500 Myr using the adaptive mesh refinement hydrodynamics code \textsc{Enzo}. We implemented a new star formation routine which deposits stars individually from 1~M$_{\odot}$ to 100~M$_{\odot}$. Using tabulated stellar properties, we follow the stellar feedback from each star. For massive stars ($M_* > 8$~M$_{\odot}$) we follow their stellar winds, ionizing radiation (using an adaptive ray-tracing radiative transfer method), the FUV radiation which leads to photoelectric heating of dust grains, Lyman-Werner radiation, which leads to H$_2$ dissociation, and core collapse supernovae. In addition, we follow the asymptotic giant branch (AGB) winds of low-mass stars ($M_* < 8$~M$_{\odot}$) and Type Ia supernovae. We investigate how this detailed model for stellar feedback drives the evolution of low mass galaxies. We find agreement with previous studies that these low mass dwarf galaxies exhibit bursty, irregular star formation histories with significant feedback-driven winds.

Using these simulations, we investigate the role that stellar radiation feedback plays in the evolution of low mass dwarf galaxies.
% a fiducial, full-physics simulation, second, a simulation with ionizing radiation restricted to the region immediately around massive star particles, and third, a simulation with no ionizing radiation feedback. I
In this regime, we find that the local effects of stellar radiation (within $\sim$ 10~pc of the massive, ionizing source star) act to regulate star formation by rapidly destroying cold, dense gas around newly formed stars. For the first time, we find that the long-range radiation effects far from the birth sites are vital for carving channels of diffuse gas in the ISM which dramatically increase the effect of supernovae. We find this effect is necessary to drive strong winds with significant mass loading factors and has a significant impact on the metal content of the ISM.
